Clusterization Properties of D-optimal Designs in Infinite Dimensional Bayesian Linear Inverse Problems

### 2023-04-13 latex/intro.tex
Need to write intro. Tell how everybody thinks model error is the solution

For the purpose of illustration, we present clusterization via a toy
inverse problem --- inferring the initial condition of the 1D heat
equation in $[0,\pi]$ with a homogeneous Dirichlet boundary condition.

The 1D heat equation is:
\begin{subequations}\label{eq:heat equation}
  \begin{alignat}{2}
    u_t &= \Delta u &&\qquad \text{in } [0,\pi] \times [0,\infty),\\
      u &= 0 &&\qquad \text{on } \{0, \pi\} \times [0,\infty),\\
        u &= u_0 &&\qquad \text{on }[0,\pi] \times \{0\}.
  \end{alignat}
\end{subequations}

Our goal here is to infer the initial condition $u_0$. We measure
$u(\x_j, T)$ at some set of locations $\x_j \in [0,\pi], j=1, \dots,m$
and a final time $T > 0$. We assume iid centered Gaussian observation
error, so we observe $v(\x_j,T) = u(\x_j,T) + \eps(\x_j)$ with
$\eps(\x_j) \sim \normal(0, \sigma^2)$ iid, $\sigma^2 > 0$. We model
the initial condition as $u_0 \sim \normal(0,\prcov)$, for $\prcov =
(-\Delta)^{-1}$ with a homogeneous Dirichlet boundary condition. It is
known \cite{Tarantola05} that for linear problems, with Gaussian prior
and error, the posterior is also Gaussian with a covariance that does
not depend on the observed data. The posterior covariance $\postcov$
has a closed form formula, even in infinite dimensions
\cite{Stuart10}. We denote by $\fwd$ the forward operator, so that $u(
\cdot,T) = \fwd u_0$, and the observation operator $\obs$ so that
$u(\x_j,T) = (\obs u)_j, j=1,\dots,m$. The posterior covariance is
known and depends only on $\prcov, \fwd, \obs$ and $\sigma^2$ (see
Section \ref{section:prelim} and \eqref{eq:postcov} specifically).

We consider the generalization of the D-optimality criterion to
infinite dimensions \cite{AlexanderianGloorGhattas14} (also, Section
\ref{subsec:D optimal design} below). We choose $\x_j,j=1,\dots,m$ to
minimize the determinant\footnote{Not exactly, see Theorem \ref{thm:}
for a precise statement.} of the posterior covariance operator.

Sensor clusterization is illustrated in Figure \ref{fig:clusterization
  illustration}. Posterior pointwise standard deviations are plotted
over the domain $[0, \pi]$. Since the posterior covariance does not
depend on data, the plot has no reference to actual data observed. The
posterior covariance does, however, depend on location of the
observation taken. Observation locations were chosen to optimize the
D-optimality criterion, and are marked by circles. Sensor
clusterization can be clearly seen for six sensors; it looks like only
four observations were taken. The reason is that two pairs of sensors
are clustered so close together that they are indistinguishable from
one another.

>>>>

For the purpose of illustration, consider a toy inverse problem ---
inferring the initial condition of the 1D heat equation in $[0,\pi]$
with a homogeneous Dirichlet boundary condition. Sensor clusterization
is clearly evident, see Figure \ref{fig:clusterization}.

The 1D heat equation with homogeneous Dirichlet boundary condition is:
\begin{subequations}\label{eq:heat equation}
  \begin{alignat}{2}
    u_t &= \Delta u &&\qquad \text{in } [0,\pi] \times [0,\infty),\\
      u &= 0 &&\qquad \text{on } \{0, \pi\} \times [0,\infty),\\
        u &= u_0 &&\qquad \text{on }[0,\pi] \times \{0\}.
  \end{alignat}
\end{subequations}

The goal is to infer the initial condition $u_0$. Model the initial
condition as $u_0 \sim \normal(0,\prcov)$, for $\prcov =
(-\Delta)^{-1}$ with a homogeneous Dirichlet boundary condition. Fix
some final time $T$ and denote by $\fwd$ the forward operator, so that
$u(\cdot, T) = \fwd u_0$. Let $\obs$ the observation operator, so that
$u(\x_j,T) = (\obs u)_j, j=1,\dots,m$. Assume iid centered Gaussian
observation error, so observations are given by $u(\x_j,T) +
\eps(\x_j)$ with $\eps(\x_j) \sim \normal(0, \sigma^2)$ iid, $\sigma^2
> 0$. The posterior covariance is known and depends only on $\prcov,
\fwd, \obs$ and $\sigma^2$ \cite{Stuart10} (see Section
\ref{section:prelim} and \eqref{eq:postcov} below). Finally, choose
$\x_j,j=1,\dots,m$ to optimize the D-optimality criterion
\cite{AlexanderianGloorGhattas14}.

Sensor clusterization is illustrated in Figure \ref{fig:clusterization
  illustration}. The posterior covariance does not depend on data, so
the plot has no reference to actual data observed. The posterior
covariance does depend on location of the observation taken, which
were chosen to optimize the D-optimality criterion. Sensor
clusterization is clearly seen for six sensors; it looks like only
four observations were taken. Two pairs of sensors are clustered so
close together that they are indistinguishable from one another.

%% A D-optimal design
%% is found via a generalization of the D-optimality criterion to
%% infinite dimension(also, Section
%% \ref{subsec:D optimal design} below).

### 2023-04-13 latex/paper.tex
Need to write intro. Tell how everybody thinks model error is the solution

\begin{abstract}
  We investigate the problem of sensor clusterization in D-optimal
  experimental design for infinite-dimensional Bayesian linear inverse
  problems. We suggest an analytically tractable model for D-optimal
  designs and reason how it may lead to sensor clusterization in the
  case of iid observation noise. We also consider how spatially
  correlated observation and model errors mitigate the clusterization
  phenomenon.

@@

\section{Acknowledgements}
This study is a part of my PhD thesis. Most of it was written under
the instruction of Prof. Georg Stadler in New York University's
Courant Institute. I would like to thank him for his great
mentorship.

>>>>

\begin{abstract}
  We investigate the problem of sensor clusterization in D-optimal
  experimental design for infinite-dimensional Bayesian linear inverse
  problems. We suggest an analytically tractable model for D-optimal
  designs. We show that sensor clusterization in the case of iid
  observation noise is generic. We also show that model spatially
  correlated errors mitigate the clusterization phenomenon.

@@

\section{Acknowledgements}
This study is a part of my PhD thesis. Most of it was written under
the instruction of Prof. Georg Stadler in the Courant Institute of
Mathematical Sciences in New York University. I would like to thank
Prof. Stadler for his great mentorship, attention to details and kind
behavior.

### 2023-04-13 latex/prelims.tex
Need to write intro. Tell how everybody thinks model error is the solution

\subsection{Overview}
In this section we will define notations that will be used throughout
the paper. Notations presented in the section are summarized in
Section \ref{subsec:notation}. The theoretical foundations for inverse
problems over function spaces can be found in \cite{Stuart10}.

\subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
Let $\hilp$ and $\hilo$ be separable Hilbert spaces (the subscripts p
and o are for ``parameter'' and ``observation'', respectfully). Denote
$\hilo^* \simeq \hilo$ the Hilbert space of all linear functionals on
$\hilo$. Consider $\fwd: \hilp \to \hilo$, a linear operator (the
``forward operator''). We are interested in forward operators that are
strongly smoothing (have fast decaying modes --- the heat operator
from section \ref{subsec:example} is a prime example). Our goal is to
infer $\param \in \hilp$ --- some parameter of the dynamics --- given
noisy observations of $\fwd \param$. We take a Gaussian prior $\pr$
for $\param$. Hence, $\param \sim \pr = \normal(\prmean ,\prcov)$ with
some appropriate covariance operator $\prcov$ \cite{Stuart10}. It is
important to note that $\fwd \prcov \fwd^*$ is the prior covariance in
$\hilo$ \cite{Stuart10}. As such, it is assumed invertible --- an
assumption we rely on below. Throughout this paper, we denote $m$ the
number of observations we are allowed to take. Observations are taken
via the operator $\obs \in ( \hilo^* )^m$. In an analogy with linear
algebra, where row vectors are thought of as linear functionals, we
think of $\obs$ as having ``rows'' $\meas_j, j=1,\dots,m$:
\begin{equation}\label{eq:O}
  \obs = (\meas_1,\dots, \meas_m)^t, \meas_j \in \hilo^*, j = 1,\dots,m.
\end{equation}
This way, for $u \in \hilo$ we have $\obs u = (\meas_1(u), \dots,
\meas_m(u) )^t \in \R^m$.
%% but we may drop the parentheses and write $\meas_j u = \meas_j(u)$.
%% A few observations regarding $\obs$ are in order.
%% First, it is good to keep in mind that $( \hilo^* )^m$ is a Hilbert
%% space with norm $\| \obs \| = \sum_{j=1}^m \|\meas_j\|$.
For $u \in \hilo$ and $v\in \R^m$:
\begin{align*}
  \big (\obs^*v \big ) (u) &= \langle v, \obs u \rangle_{\R^m} = \sum_{j=1}^m  v_j \meas_j(u)
  = v^t \left ( \obs u \right ) = (v^t \obs) (u),
\end{align*}
and thus:
\begin{align}\label{eq:obs*}
  \obs^*v &= \sum_{j=1}^m v_j \meas_j = v^t \obs.
\end{align}

@@

Each observation is a linear functional, chosen from some subset of
$\hilo^{*}$. Data is acquired via noisy observations
\begin{align}\label{eq:inverse problem}
  \data := \obs (\fwd \param + \eps') + \eps = \obs \fwd \param + \obs \eps' + \eps,
\end{align}
with $\data \in \R^m$ and $\eps, \eps'$ defined next. We consider two
types of error terms. First, there is spatially correlated model error
$\eps' \sim \normal(0,\modcov)$, modeled as a centered Gaussian
measure on $\hilo$ with covariance operator $\modcov$. Then there is
observation error $\eps \sim \normal(0, \sigma^2 I_m)$, with $I_m$ the
$m \times m$ identity. Both error terms and the prior are assumed
independent of each other. In the literature, it is common for the
observation and forward operators to be merged $\tmp := \obs \fwd$
\cite{AlexanderianGloorGhattas14}. Since we need to separate the
observation operator $\obs$ from the forward operator $\fwd$ in order
to carry out the analysis in the following sections, we explicitly
separate $\fwd$ and $\obs$ in \eqref{eq:inverse problem}. We now turn
to understand the terms present in \eqref{eq:inverse problem}.

\subsubsection{Dynamics and Observation Operators}\label{subsec:dynamics}
Rows of $\obs$ are linear functionals and we refer to these as {\it
  observation vector}s. Note that we cannot choose every observation
vector we want. For example, we may be restricted only to pointwise
evaluations of $\fwd \param$ by the sensors at our disposal, in which
case measuring, e.g., the mean $\ell(u) = \int_{\Omega}u$ of a
function is not possible. Thus, it is best to think of $(\obs \fwd
\param)_j$ as approximations to pointwise evaluations. For example, if
$\hilo = L^2([-\pi, \pi])$, consider $\ell_1(x) := (2\pi
\eps)^{-\frac{1}{2}}\exp(-\frac{x^2}{2\eps^2})$ and $\ell_2(x)
:=\eps^{-1}\mathbf{1}_{[-\frac{\eps}{2}, \frac{\eps}{2}]}$. Both
$\ell_1$ and $\ell_2$ approximate a point measurement at $0$ for small
$\eps$ and, crucially, $\ell_1, \ell_2 \in \hilo^*$.
%% It is important to note that although $\|\ell\|_1 =
%% \|\ell_2\|_1=1$, unfortunately $\|\ell_1\|_2^2 = (4\pi
%% \eps)^{-\frac{1}{2}}$ and $\|\ell_2\|_2^2 = \eps^{-1}$.

\subsubsection{Error Terms}
It is easy to verify that $\bar{\eps} := \obs \eps' + \eps \in \R^m$
is a centered Gaussian random vector with covariance
matrix

\begin{align}\label{eq:Sigma}
  \begin{split}
    \Sigma(\obs) :&= \mathbb{E}[ (\obs \eps' + \eps) (\obs \eps' +
      \eps)^t ]
    % 
    % 
    = \obs \modcov \obs^* + \sigma^2I , 
  \end{split}
\end{align}
where
\begin{equation}\label{eq:modcov explained}
  [\obs \modcov \obs^*]_{ij} = e_i^t \obs \modcov \obs^* e_j = \meas_i (\modcov \meas_j),
\end{equation}
and $e_j \in \R^m$ is the $j$th standard basis vector. The first
equality in \eqref{eq:modcov explained} holds by definition and the
second by \eqref{eq:obs*}. The explicit dependence on $\obs$ will be
mostly dropped for notational convenience, so $\Sigma(\obs) =
\Sigma$. Thus, for a fixed $\obs$ (i.e.\ a fixed set of observations)
we can equivalently write $\data = \obs \fwd \param + \bar{\eps}$ with
$\bar{\eps} \sim \normal(0,\Sigma)$. Taking $\modcov = 0$ is common
practice \cite{Tarantola05,Kaipio2006,Vogel02}. This reduces to the
case where we take iid observations. Then $\Sigma = \sigma^2I$ is
simply a scalar matrix that does not depend on $\obs$. Note that
taking an error model with a non-scalar covariance as we do here
allows us to consider model error (modeled by $\modcov$) as well as
observation error (modeled by $\sigma^2$). For example, say we believe
our forward model does not capture some small scale phenomenon.  Then
we may express this belief by saying $\tru = \fwd + \err$, with $\fwd$
depending on $\param$ and $\err$ depending on $\sspar$, with $\param
\perp \sspar$. We do not know much about this effect but it is
reasonable to assume it changes continuously in our domain. We (may
choose to) model it as $\normal (0, \modcov)$ and take $\modcov$ to
reflect the spatial (or other) variability we imagine $\err \sspar$
has. Such small scale phenomenon can arise as a modeling issue, where
we might not model the system in its entirety. It can also arise from
a numerical source, where our discretization of the system is not fine
enough to capture all small scale phenomena.
%%  In section \ref{section:non vanishing}, we will see that assuming
%% some correlation in the error mitigates the clusterization
%% phenomenon, as reported in the literature \cite{Ucinski05}.

Finally, it is useful to record a known result regarding the posterior
covariance operator of our inverse problem:
\begin{align}\label{eq:postcov}
  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd
  )^{-1}.
\end{align}

@@

The meaning of D-optimal design in infinite-dimensional Hilbert spaces
was investigated in \cite{AlexanderianGloorGhattas14}. There, the
authors make assumptions that amount to $\Sigma=I$ (implied by
$\modcov = 0,\sigma^2=1$), but we choose not to take these
simplification here. This is because $\Sigma$ can determine ``how
much'' clusterization we see. As stated
\cite[pp. 681]{AlexanderianGloorGhattas14}, the results hold for more
general covariance matrices. The conclusion is that in
infinite-dimensions, a D-optimal design is well-defined as maximizing
the expected KL divergence between posterior and prior. The main
result is summarized, using our notation, in the following theorem:
\begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]\label{thm:d optimality}
  Let $\pr = \normal(\prmean,\prcov)$ be a Gaussian prior on $\hilp$
  and let $\post = \normal(\postmean,\postcov)$ the posterior measure
  on $\hilp$ for the Bayesian linear inverse problem $\data = \obs
  \fwd\param + \obs \eps' + \eps$ discussed above. Then
  \begin{align}\label{eq:objective}
    \begin{split}
      \tar( \obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
      % 
      % 
      % 
      &= \frac12 \log \det 
      ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}).
    \end{split}
  \end{align}
\end{theorem}
\begin{definition}\label{def:d optimality}
  A design $\obs^{\star}$ is said to be D-optimal if $\obs^{\star} = \argmax_{\obs} \tar(\obs)$.
\end{definition}

\noindent The intuition behind Definition \ref{def:d optimality} is
straightforward. From \eqref{eq:postcov} and \eqref{eq:objective}:
\begin{align*}
  \begin{split}
    \det ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}) &= \det \Big( \prcov ( \prcov^{-1} + \fwd ^* \obs^* \Sigma^{-1} \obs \fwd) \Big )\\
    &= \det \prcov \det \postcov^{-1}.
  \end{split}
\end{align*}
Since $\prcov$ is constant, a D-optimal design minimizes the posterior
covariance determinant, analogously to the finite-dimensional case.

>>>>

\subsection{Overview}
In this section I will present notations that will be used throughout
this article. These notations are summarized in Section
\ref{subsec:notation}. The theoretical foundations for inverse
problems over function spaces can be found in \cite{Stuart10}.

\subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
Let $\hilp$ and $\hilo$ be separable Hilbert spaces (the subscripts p
and o are for ``parameter'' and ``observation'', respectfully). Denote
$\hilo^* \simeq \hilo$ the Hilbert space of all linear functionals on
$\hilo$. Let $\fwd: \hilp \to \hilo$ the \emph{forward operator}. The
forward operator $\fwd$ is assumed linear and strongly smoothing (has
fast decaying modes --- the heat operator from section
\ref{subsec:example} is a prime example). Take a Gaussian prior
$\param \sim \pr = \normal(\prmean ,\prcov)$ with some appropriate
covariance operator $\prcov$ \cite{Stuart10} on $\hilp$. Note that
$\fwd \prcov \fwd^*$ is the prior covariance in $\hilo$
\cite{Stuart10}. As such, it is assumed invertible --- an assumption
which will be used heavily below. Let $m$ the number of observations
taken. Observations are taken via the \emph{observation
operator}\footnote{It is common for the observation and forward
operators to be merged $\tmp := \obs \fwd$
\cite{AlexanderianGloorGhattas14}, but the analysis carried out in the
following sections requires $\fwd$ and $\obs$ are explicitly
separated.} $\obs \in ( \hilo^* )^m$. In an analogy with linear
algebra, where row vectors are thought of as linear functionals,
$\obs$ is best thought of as having ``rows'' $\meas_j, j=1,\dots,m$:
\begin{equation}\label{eq:O}
  \obs = (\meas_1,\dots, \meas_m)^t, \meas_j \in \hilo^*, j = 1,\dots,m.
\end{equation}
This way, $\obs u = (\meas_1(u), \dots, \meas_m(u) )^t \in \R^m$ and
$u \in \hilo$.
%% but we may drop the parentheses and write $\meas_j u = \meas_j(u)$.
%% A few observations regarding $\obs$ are in order.
%% First, it is good to keep in mind that $( \hilo^* )^m$ is a Hilbert
%% space with norm $\| \obs \| = \sum_{j=1}^m \|\meas_j\|$.
For $u \in \hilo$ and $v\in \R^m$:
\begin{align*}
  \big (\obs^*v \big ) (u) &= \langle v, \obs u \rangle_{\R^m} = \sum_{j=1}^m  v_j \meas_j(u)
  = v^t \left ( \obs u \right ) = (v^t \obs) (u),
\end{align*}
and thus:
\begin{align}\label{eq:obs*}
  \obs^*v &= \sum_{j=1}^m v_j \meas_j = v^t \obs.
\end{align}

@@

Each observation vector is a linear functional, chosen from (some
subset of) $\hilo^{*}$. Data is acquired via noisy observations, with
two types of error terms considered. Spatially correlated model error
is denoted $\eps' \sim \normal(0,\modcov)$, and modeled as a centered
Gaussian measure on $\hilo$ with covariance operator
$\modcov$. Observation error is denoted $\eps \sim \normal(0, \sigma^2
I_m)$, with $I_m \in \mathbb{R}^{m \times m}$ the identity. Both error
terms and the prior are assumed independent of each other. Thus, data
is acquired via
\begin{align}\label{eq:inverse problem}
  \data := \obs (\fwd \param + \eps') + \eps = \obs \fwd \param + \obs \eps' + \eps.
\end{align}
%% \subsubsection{Error Terms}
It is easy to verify that $\obs \eps' + \eps \in \R^m$ is a centered
Gaussian random vector with covariance matrix

\begin{align}\label{eq:Sigma}
  \begin{split}
    \Sigma(\obs) :&= \mathbb{E}[ (\obs \eps' + \eps) (\obs \eps' +
      \eps)^t ]
    % 
    % 
    = \obs \modcov \obs^* + \sigma^2I , 
  \end{split}
\end{align}
where
\begin{align}\label{eq:modcov explained}
  \begin{split}
    [\obs \modcov \obs^*]_{ij} & = e_i^t \obs \modcov \obs^* e_j \\
    %
    %
    %
    &= \meas_i (\modcov \meas_j) \text{ (by \eqref{eq:obs*})}.
  \end{split}
\end{align}
%% and $e_j \in \R^m$ is the $j$th standard basis vector.
%% The explicit dependence on $\obs$ will be mostly dropped for
%% convenience, so $\Sigma(\obs) = \Sigma$.  Thus, for a fixed $\obs$
%% (i.e.\ a fixed set of observations), $\data = \obs \fwd \param +
%% \bar{\eps}$ with $\bar{\eps} \sim \normal(0,\Sigma)$.
Taking $\modcov = 0$ is a common practice
\cite{Tarantola05,Kaipio2006,Vogel02} and then $\Sigma = \sigma^2I$ is
simply a scalar matrix that does not depend on $\obs$.

%% Taking an error model with a non-scalar covariance allows us to
%% consider model error (modeled by $\modcov$) as well as observation
%% error (modeled by $\sigma^2$). For example, say we believe our forward
%% model does not capture some small scale phenomenon.  Then we may
%% express this belief by saying $\tru = \fwd + \err$, with $\fwd$
%% depending on $\param$ and $\err$ depending on $\sspar$, with $\param
%% \perp \sspar$. We do not know much about this effect but it is
%% reasonable to assume it changes continuously in our domain. We (may
%% choose to) model it as $\normal (0, \modcov)$ and take $\modcov$ to
%% reflect the spatial (or other) variability we imagine $\err \sspar$
%% has. Such small scale phenomenon can arise as a modeling issue, where
%% we might not model the system in its entirety. It can also arise from
%% a numerical source, where our discretization of the system is not fine
%% enough to capture all small scale phenomena.
%%  In section \ref{section:non vanishing}, we will see that assuming
%% some correlation in the error mitigates the clusterization
%% phenomenon, as reported in the literature \cite{Ucinski05}.

%% \subsubsection{Dynamics and Observation Operators}\label{subsec:dynamics}
%% Rows of $\obs$ are linear functionals and we refer to these as {\it
%%   observation vector}s. Note that we cannot choose every observation
%% vector we want. For example, we may be restricted only to pointwise
%% evaluations of $\fwd \param$ by the sensors at our disposal, in which
%% case measuring, e.g., the mean $\ell(u) = \int_{\Omega}u$ of a
%% function is not possible. Thus, it is best to think of $(\obs \fwd
%% \param)_j$ as approximations to pointwise evaluations. For example, if
%% $\hilo = L^2([-\pi, \pi])$, consider $\ell_1(x) := (2\pi
%% \eps)^{-\frac{1}{2}}\exp(-\frac{x^2}{2\eps^2})$ and $\ell_2(x)
%% :=\eps^{-1}\mathbf{1}_{[-\frac{\eps}{2}, \frac{\eps}{2}]}$. Both
%% $\ell_1$ and $\ell_2$ approximate a point measurement at $0$ for small
%% $\eps$ and, crucially, $\ell_1, \ell_2 \in \hilo^*$.
%% %% It is important to note that although $\|\ell\|_1 =
%% %% \|\ell_2\|_1=1$, unfortunately $\|\ell_1\|_2^2 = (4\pi
%% %% \eps)^{-\frac{1}{2}}$ and $\|\ell_2\|_2^2 = \eps^{-1}$.

Finally, it is useful to record a known result regarding the posterior
covariance operator:
\begin{align}\label{eq:postcov}
  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd
  )^{-1}.
\end{align}

@@

D-optimal designs in infinite-dimensional Hilbert spaces were
investigated in \cite{AlexanderianGloorGhattas14}. In
infinite-dimensions, a D-optimal design is well-defined as maximizing
the expected KL divergence between posterior and prior. The main
result is summarized, in the notation presented above, in the
following theorem\footnote{In \cite{AlexanderianGloorGhattas14},
results are stated for $\Sigma=I$ (implied by $\modcov =
0,\sigma^2=1$), but the authors also mention these results hold for
more general covariance matrices (p. 681).}:
\begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]\label{thm:d optimality}
  Let $\pr = \normal(\prmean,\prcov)$ be a Gaussian prior on $\hilp$
  and let $\post = \normal(\postmean,\postcov)$ the posterior measure
  on $\hilp$ for the Bayesian linear inverse problem $\data = \obs
  \fwd\param + \obs \eps' + \eps$ discussed above. Then
  \begin{align}\label{eq:objective}
    \begin{split}
      \tar( \obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
      % 
      % 
      % 
      &= \frac12 \log \det 
      ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}).
    \end{split}
  \end{align}
\end{theorem}
\begin{definition}\label{def:d optimality}
  $\obs^{\star}$ is said to be \emph{D-optimal} if $\obs^{\star} =
  \argmax_{\obs} \tar(\obs)$, where rows of $\obs$ are in some allowed
  set of measurement.
\end{definition}

\noindent In order to undestand the intuition behind Definition
\ref{def:d optimality}, note that from \eqref{eq:postcov} and
\eqref{eq:objective}:
\begin{align*}
  \begin{split}
    \tar(\obs) &= \det ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}) \\
    &= \det \Big( \prcov ( \prcov^{-1} + \fwd ^* \obs^* \Sigma^{-1} \obs \fwd) \Big )\\
    &= \det \prcov \det \postcov^{-1}.
  \end{split}
\end{align*}
Since $\prcov$ is constant, a D-optimal design minimizes the posterior
covariance determinant, analogously to the finite-dimensional case.
